---
title: "Tidy Tuesday Art Collections"
author: "Jeff Grayum"
date: "1/14/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Libraries:
```{r}
library(tidyverse)
library(tidytuesdayR)
library(ggthemes)
library(scales)
library(tidytext)
library(tidymodels)
#install.packages("textrecipes")
library(textrecipes)
#install.packages("hardhat")
library(hardhat)
library(doParallel)
#install.packages("glmnet")
library(glmnet)
library(janitor)
```

Loading our data:
```{r}
tuesdata <- tidytuesdayR::tt_load('2021-01-12')

artists <- tuesdata$artists 
  

artwork <- tuesdata$artwork %>%
  filter(artist != "Turner, Joseph Mallord William")
```


Exploring the data:
```{r}
artists %>% 
  view()

artists %>%
  count(name, sort = TRUE)

artists %>%
  count(gender, sort = TRUE)

artists %>%
  filter(!is.na(yearOfBirth),
         !is.na(yearOfDeath)) %>%
  mutate(age = yearOfDeath - yearOfBirth) %>%
  filter(age >= 1) %>%
  ggplot(aes(age, fill = gender)) +
  geom_histogram(binwidth = 5) +
  theme_fivethirtyeight() +
  labs(x = "Age", 
       y = "",
       fill = "",
       title = "Age and gender of historical artists") +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

artists %>%
  filter(!is.na(yearOfBirth)) %>%
  mutate(birth_decade = (yearOfBirth %/% 10) * 10) %>%
  ggplot(aes(birth_decade, fill = gender)) +
  geom_bar() +
  labs(x = "Decade of birth",
       y = "",
       title = "When were famous artist's born?") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

artists %>%
  filter(!is.na(yearOfBirth),
                  !is.na(placeOfBirth)) %>%
  mutate(birth_decade = (yearOfBirth %/% 10) * 10) %>%
  count(placeOfBirth, birth_decade, sort = TRUE) %>%
  arrange(birth_decade) %>%
  filter(n >= 5) %>%
  mutate(fct_reorder(placeOfBirth, n, sum)) %>%
  ggplot(aes(birth_decade, n, fill = placeOfBirth)) +
  geom_col() +
  theme_fivethirtyeight() +
  labs(x = "Birth decade",
       y = "",
       fill = "Place of birth",
       title = "Artist hot spots throughout time") +
  theme(axis.title = element_text(),
  text = element_text(family = "Times New Roman"))

artists %>%
  filter(!is.na(placeOfBirth)) %>%
  mutate(placeOfBirth = fct_lump_n(placeOfBirth, 18)) %>%
  count(placeOfBirth, sort = TRUE) %>%
  mutate(placeOfBirth = fct_reorder(placeOfBirth, n)) %>%
  ggplot(aes(n, placeOfBirth, fill = placeOfBirth)) +
  geom_col() +
  labs(x= "",
       y = "",
       title = "Birthplace of famous artists") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"),
        legend.position = "none")

artists %>%
  filter(!is.na(yearOfBirth),
         !is.na(gender)) %>%
  mutate(birth_decade = (yearOfBirth %/% 10) * 10) %>%
  count(birth_decade, gender, sort = TRUE) %>%
  ggplot(aes(birth_decade, n, color = gender)) +
  geom_line(size = 1) +
  labs(x = "Year of birth",
       y = "",
       title = "Gender differences among artists througout time",
       color = "") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))
  
```


```{r}
artwork %>%
  view()

artwork %>%
  mutate(artist = fct_lump(artist, 10)) %>%
  add_count(artist) %>%
  mutate(artist = fct_reorder(artist, n)) %>%
  ggplot(aes(artist, n, fill = artistRole)) +
  geom_col()

artwork %>%
  count(artist, artistRole, sort = TRUE) %>%
  head(10) %>%
  mutate(artist = fct_reorder(artist, n)) %>%
  ggplot(aes(n, artist, fill = artistRole)) +
  geom_col() +
  labs( x = "",
        y = "Artist",
        title = "Most prolific artists in history",
        fill = "Role") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

artwork %>%
  group_by(artist) %>%
  summarize(n_pieces = n()) %>%
  mutate(artist = fct_reorder(artist, n_pieces)) %>%
  head(10) %>%
  view()

artwork %>%
  count(artist, sort = TRUE)
  
```

Now, for Julia's analysis!  Training a linearized model with text data to show the relationship between the medium of the artwork and how that changes over time.
```{r}
artwork %>%
  glimpse()

artwork %>%
  count(medium, sort = TRUE)

artwork %>%
  ggplot(aes(year)) +
  geom_histogram(alpha = 0.8, fill = "midnightblue") +
  labs(x = "Year",
       y = "",
       title = "Distribution of art pieces") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))  #bimodal distribution

tate_df <- artwork %>%
  filter(year >= 1750) %>%
  select(year, medium) %>%
  na.omit() %>%
  arrange(year)

#Let's see what the most common words in artwork$medium are.

tate_df %>%
  unnest_tokens(word, medium) %>%
  count(word, sort = TRUE)
```


Let's build our model, using tidymodels.  Set seed, split, train, test.
```{r}
set.seed(123)

art_split <- initial_split(tate_df, strata = year)
art_train <- training(art_split)
art_test <- testing(art_split)


#Now we will resample the training data.
set.seed(234)
art_folds <- vfold_cv(art_train, strata = year)
art_folds
```

First, some pre-processing.
```{r}
#Explain year, features coming from medium.  Learn from training data, apply to testing data and new data.

#pre-processing recipe for this dataset of text.
art_rec <- recipe(year ~ medium, data = art_train) %>%
  step_tokenize(medium) %>%
  step_stopwords(medium) %>%
  step_tokenfilter(medium, max_tokens = 500) %>%
  step_tfidf(medium)

art_rec
```


```{r}
sparse_bp <- hardhat::default_recipe_blueprint(composition = "dgCMatrix")

lasso_spec <- linear_reg(penalty = tune(), mixture = 1)  %>%
  set_engine("glmnet") 

art_wf <- workflow() %>%
  add_recipe(art_rec, blueprint = sparse_bp) %>%
  add_model(lasso_spec)

art_wf
```

```{r}
doParallel::registerDoParallel()

lambda_grid <- grid_regular(penalty(range = c(-3, 0)), levels = 20)

lasso_rs <- tune_grid(
  art_wf,
  resamples = art_folds,
  grid = lambda_grid
)

lasso_rs
```


```{r}
artwork %>% 
  view()

artwork %>%
  glimpse()

artwork <- artwork %>%
  clean_names() 

artists <- artists %>%
  clean_names()

```

```{r}
artwork %>%
  count(medium, sort = TRUE) %>%
  head(50) %>%
  mutate(medium = fct_reorder(medium, n)) %>%
  ggplot(aes(n, medium)) +
  geom_col() +
  labs(x = "",
       y = "Medium",
       title = "Most post popular mediums for art") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

artwork %>%
  extract(medium, c("medium1", "medium2"), "(.*) on (.*)") %>%
  count(medium1, sort = TRUE) %>%
  head(50) %>%
  mutate(medium1 = fct_reorder(medium1, n)) %>%
  ggplot(aes(n, medium1)) +
  geom_col() +
  scale_x_log10() +
  labs(x = "",
       y = "Medium",
       title = "Most post popular mediums for art") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

artwork %>%
  extract(medium, c("medium1", "medium2"), "(.*) on (.*)") %>%
  count(medium2, sort = TRUE) %>%
  head(50) %>%
  mutate(medium2 = fct_reorder(medium2, n)) %>%
  ggplot(aes(n, medium2)) +
  geom_col() +
  labs(x = "",
       y = "Medium",
       title = "Most post popular mediums for art") +
  theme_fivethirtyeight() +
  scale_x_log10() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

artwork %>%
  separate(medium, c("medium1", "surface"), sep = " on " ,
           fill = "right", extra = "merge") %>%
  count(medium1, sort = TRUE) %>%
  head(20) %>%
  replace_na(list(medium1 = "Missing")) %>%
  mutate(medium1 = fct_reorder(medium1, n)) %>%
  ggplot(aes(n, medium1)) +
  geom_col() +
  labs(x = "",
       y = "Medium",
       title = "Most post popular art mediums in the Tate Museum") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

#Remember this cool trick that let's use filter rather than fct_lump?
artwork %>%
  filter(fct_lump(artist, 16) != "Other") %>%
  mutate(medium = fct_lump(medium, 6)) %>%
  count(artist, medium, sort = TRUE) %>%
  mutate(artist = fct_reorder(artist, n, sum)) %>%
  ggplot(aes(n, artist, fill = medium)) +
  geom_col() +
  labs(x = "",
       y = "Artist",
       title = "Most post popular artists in the Tate Museum",
       fill = "Medium") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))
  
```


Let's look at how medium's change over time. Cool use of count(), and complete!
```{r}
#First, look at the sum of medium's over time

artwork %>%
  separate(medium, c("medium", "surface"), sep = " on " ,
           fill = "right", extra = "merge") %>%
  filter(year >= 1750) %>%
  count(decade = round(year, -1),
        medium = fct_lump(medium, 8)) %>%
  complete(decade, medium, fill = list(n = 0)) %>%
  ggplot(aes(decade, n, fill = medium)) +
  geom_area() +
  labs(x = "Decade",
       y = "",
       fill = "Medium",
       title = "Changes in art mediums over time") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

#Now, let's look at the percent of medium's over time.

by_decade_medium <- artwork %>%
  separate(medium, c("medium", "surface"), sep = " on " ,
           fill = "right", extra = "merge") %>%
  filter(year >= 1750) %>%
  count(decade = round(year, -1),
        medium = fct_lump(medium, 11)) %>%
  complete(decade, medium, fill = list(n = 0)) %>%
  group_by(decade) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() 

by_decade_medium %>%
  filter(!is.na(medium)) %>%
  mutate(medium = fct_reorder(medium, n, sum)) %>%
  ggplot(aes(decade, pct, fill = medium)) +
  geom_area() +
  scale_y_continuous(labels = percent) +
  labs(x = "Decade",
       y = "Percent of Tate Museum's pieces",
       fill = "Medium",
       title = "Changes in art mediums over time") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

#Let's visualize these data in another way!
by_decade_medium %>%
  filter(!is.na(medium)) %>%
  mutate(medium = fct_reorder(medium, pct, sum, .desc = TRUE)) %>%
  ggplot(aes(decade, pct, fill = medium)) +
  geom_col() +
  facet_wrap(~ medium, scales = "free_y") +
  scale_y_continuous(labels = percent) +
  labs(x = "Decade",
       y = "Percent of Tate Museum's pieces",
       fill = "Medium",
       title = "Changes in art mediums over time") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"),
        legend.position = "none")

by_decade_medium <- artwork %>%
  separate(medium, c("medium", "surface"), sep = " on " ,
           fill = "right", extra = "merge") %>%
  filter(year >= 1750) %>%
  group_by(decade = round(year, -1)) %>%
  mutate(n_artists_decade = n_distinct(artist)) %>%
  ungroup() %>%
  group_by(decade,
        medium = fct_lump(medium, 11),
        n_artists_decade) %>%
  summarize(n = n(),
            n_artists = n_distinct(artist)) %>%
  complete(decade, medium, fill = list(n = 0, n_artists = 0)) %>%
  group_by(decade) %>%
  mutate(pct = n / sum(n),
         pct_artists = n_artists / n_artists_decade) %>%
  ungroup() 

by_decade_medium %>%
  filter(!is.na(medium)) %>%
  mutate(medium = fct_reorder(medium, pct_artists, sum, .desc = TRUE)) %>%
  ggplot(aes(decade, pct_artists, fill = medium)) +
  geom_col() +
  facet_wrap(~ medium, scales = "free_y") +
  scale_y_continuous(labels = percent) +
  labs(x = "Decade",
       y = "Percent of artists using medium",
       fill = "Medium",
       title = "Changes in art mediums over time") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"),
        legend.position = "none")

```

